```{r, include=F, purl = F}
# knitr::opts_knit$set(root.dir = "/home/rata/ResearchMethods/14_Statistik2/") 
```


```{r, message=F, include=F}
library(tidyverse)
```


#### Beispiel 2: Spearman rank correlation

[Script als Download](14_Statistik2/RFiles/Demo_01_Korr_Reg_Beispiel2.R)


Weihnachtsinsel-Krabben aus @logan2010 [S. 186] (Script an Tools/Packages des Moduls Research Methods angepasst)

Green (1997) investigated the correlation between total biomass of red land crabs (*Gecarcoidea natalis*) and the density of their burrows at a number of forested sites (Lower site: LS and Drumsite: DS) on Christmas Island.

##### Step 1

Import the Green (1997) data set [green.csv](14_Statistik2/data/green.csv)

```{r, message = F}

green <- read_csv("14_Statistik2/data/green.csv")

```


##### Step 2

Assess linearity and bivariate normality for the two sites separately using a scatterplots with marginal boxplots

```{r, fig.cap="some evidence of non-normality (boxplots not asymmetrical)"}


p1 <- green %>%
  filter(SITE == "LS") %>%
  ggplot(aes(TOTMASS,BURROWS)) +
  geom_point() +
  geom_smooth(se = F)

p2 <- green %>%
  filter(SITE == "DS") %>%
  ggplot(aes(TOTMASS,BURROWS)) +
  geom_point() +
  geom_smooth(se = F)

ggMarginal(p1, type = "boxplot",size = 15)
```

```{r}
ggMarginal(p2, type = "boxplot",size = 15)

```

##### Step 3

Calculate the Spearman’s rank correlation coefficient and test $H_0: \rho = 0$ (that the population correlation coefficient equals zero)

```{r}
cor.test(~BURROWS + TOTMASS, data = green, subset = SITE == "LS", method="spearman")
```

Conclusions - reject $H_0$ that population correlation coefficient equals zero, there was a strong positive correlation between crab biomass and burrow density at Low site ($\rho = 0.851$, $S_{10} = 24.57$, $P = 0.0018$).

```{r}
cor.test(~BURROWS + TOTMASS, data = green, subset = SITE == "DS",method="spearman")
```

Conclusions - do not reject $H_0$ that population correlation coefficient equals zero, there was no detectable correlation between crab weight and gill weight at Drumsite ($\rho = 0.168$, $S_{10} = 69.92$, $P = 0.692$).

##### Step 4

Summarize findings with scatterplots, including a 95% confidence ellipses for the population bivariate mean center.


```{r}
ggplot(green, aes(BURROWS, TOTMASS)) +
  geom_point() +
  stat_ellipse(geom = "path", level = 0.95) +
  facet_wrap(~SITE)
```