```{r, echo = F, purl = F}
knitr::opts_chunk$set(echo = T, collapse=TRUE)
# knitr::opts_knit$set(root.dir = "13_Statistik1") 
library(knitr)

```





## Demo: Stastische Tests

[Demoscript als Download](13_Statistik1/RFiles/Demo_Tests.R)

### Assoziationstests

```{r, message = F}
library(car)
library(tidyverse)

library(ggfortify)

select <- dplyr::select

```




```{r}

# Chi-Quadrat-Test

# Ermitteln des kritischen Wertes für 95 perzentile und 1 FG
qchisq(0.95,1)


```

```{r}

# Datensatz von Folie 28: Test auf Assoziation zwischen zwei kategorialen Variablen
# Frage: Wie hängen zwei Eigenschaften des gleichen Objektes zusammen?

count <- matrix(c(38,14,11,51),nrow=2)

count

dimnames(count) <- list(c("helle_haare","hunkle_haare"),c("blaue_augen","braune_augen"))

count

chisq.test(count)

chisq.test(count,correct = F)

fisher.test(count)
```




###	t-Test


```{r}

# Datensatz "Blumen" erstellen für t-Test

blume <-data.frame(
  sorte_a = c(20,19,25,10,8,15,13,18,11,14),
  sorte_b = c(12,15,16,7,8,10,12,11,13,10)
  )

blume
summary(blume)

```


#### Tests mit einer *breiten* (wide) Tabelle
```{r}



boxplot(blume$sorte_a,blume$sorte_b)

hist(blume$sorte_a)

hist(blume$sorte_b)

t.test(blume$sorte_a,blume$sorte_b) #zweiseitig
t.test(blume$sorte_a,blume$sorte_b, alternative="greater") #einseitig
t.test(blume$sorte_a,blume$sorte_b, alternative="less") #einseitig
t.test(blume$sorte_a,blume$sorte_b, var.equal=T) #Varianzen gleich, klassischer t-Test
t.test(blume$sorte_a,blume$sorte_b, var.equal=F) #Varianzen ungleich, Welch's t-Test, ist auch default
t.test(blume$sorte_a,blume$sorte_b, paired=T) #gepaarter t-Test 
t.test(blume$sorte_a,blume$sorte_b, paired=T,alternative="greater") #gepaarter t-Test 

shapiro.test(blume$sorte_b)

wilcox.test(blume$sorte_a,blume$sorte_b)

```


#### Tests mit einer *langen* (long) Tabelle
```{r}
# T-Tests mit einer langen (long) Tabelle

# breit zu long mit tidyr::gather

blume_long <- gather(blume,sorte,groesse)

# Für eine Long table können wir auch gut ggplot verwenden

ggplot(blume_long, aes(sorte, groesse)) +
  geom_boxplot()

ggplot(blume_long, aes(groesse)) +
  geom_histogram(binwidth = 1, colour = "white")

ggplot(blume_long, aes(groesse)) +
  geom_histogram(binwidth = 2, colour = "white")+
  facet_grid(sorte~.)

ggplot(blume_long, aes(groesse, fill = sorte)) +
  geom_density(alpha = 0.5)




t.test(groesse~sorte, blume_long)

# Optionen "alternative" und "var.equal" werden angepasst analog der "wide" table

```





### Randomisierung

Datensatz: [beetle.csv](13_Statistik1/data/beetle.csv)

```{r}
# Randomisierung

beetles <- read_delim("13_Statistik1/data/beetle.csv", ",")


ggplot(beetles, aes(SIZE, BEETLES)) +
  geom_boxplot()


ggplot(beetles, aes(SIZE, sqrt(BEETLES))) +
  geom_boxplot()


ggplot(beetles, aes(SIZE, BEETLES)) +
  geom_boxplot() +
  scale_y_sqrt()


# extract specific inidies from t-test
stat <- function(data, indices) {
  t.test <- t.test(BEETLES~SIZE, data)$"stat"
  t.test
}


# Random generator auf der Basis eines Datensatzes
rand.gen <- function(data,mle) {
  out <- data
  out$SIZE <- sample(out$SIZE, replace=F)
  out
}


library(boot)

# boot(): bootstrap resampling

beetles.boot <- boot(data = beetles,
                     statistic = stat, 
                     R=5000, 
                     sim="parametric", 
                     ran.gen=rand.gen)


print(beetles.boot)



plot(beetles.boot)



tval <- length(beetles.boot[beetles.boot$t >= abs(beetles.boot$t0)])+1

tval

tval/(beetles.boot$R + 1)

```


###	Voraussetzungen parametrischer Verfahren


#### F-Test
```{r}

# F-Test

var.test(blume$sorte_a,blume$sorte_b)
var.test(groesse~sorte, blume_long)

```


#### Levene Test
```{r}

# Levene Test 

library(car)
leveneTest(blume$sorte_a,blume$sorte_b,center=mean)
leveneTest(groesse~sorte, blume_long)

```

#### Visuelle Inspektion / Transformation

```{r}

# Visuelle Inspektion

sleep <- msleep %>%
  filter(vore %in% c("carni","herbi")) %>%
  dplyr::select(name,vore,bodywt)


ggplot(sleep, aes(vore, bodywt, group = vore)) +
  geom_boxplot()


sleep <- sleep %>%
  mutate(
    bodywt_log10 = log10(bodywt),
    bodywt_sqrt = sqrt(bodywt),
    bodywt_4throot = bodywt^0.25
  ) %>%
  gather(key,val, -c(name,vore))

sleep


ggplot(sleep, aes(vore, val, group = vore)) +
  geom_boxplot() +
  facet_wrap(~key, scales = "free_y")




```


###	Einstieg ANOVA

#### ANOVA mit "Blumen"-Daten

```{r}

# ANOVA mit "Blumen"-Daten

ggplot(blume_long, aes(sorte, groesse)) +
  geom_boxplot()

t.test(groesse~sorte, blume_long, var.equal=T)

#now as ANOVA

aov(groesse~sorte,blume_long)
summary(aov(groesse~sorte,blume_long)) #F-value = 4.325
summary.lm(aov(groesse~sorte,blume_long))

lm(groesse~sorte,blume_long)
summary(lm(groesse~sorte,blume_long))

# Compare the above results of t.test, aov and lm and how the relevant data are displayed

# Now check with the F-distribution:
# What would have been the critical value for a significant result at p < 0.05?

qf(0.95,1,18)

# which is the p-value associated with the obtained F-value?

1-pf(4.325,1,18) #probability of obtaining F=4,325 or greater

```


```{r}
## Optional für Demo

#overall mean and residuals

blume_long$index <- 1:20

blume_long <- blume_long %>%
  group_by(sorte) %>%
  mutate(
    mean = mean(groesse)
  )

ggplot(blume_long, aes(index,groesse)) +
  geom_point() +
  geom_hline(aes(yintercept =  mean(groesse))) +
  geom_segment(aes(x = index,y = groesse,xend = index,yend = mean(groesse))) +
  labs(x = "Order",y = expression(Size~(cm^2)))

#group means and residuals

ggplot(blume_long, aes(index,groesse, colour = sorte)) +
  geom_point() +
  geom_hline(aes(yintercept = mean)) +
  geom_segment(aes(x = index,y = groesse,xend = index,yend = mean)) +
  labs(x = "Order",y = expression(Size~(cm^2))) +
  facet_grid(~sorte)



library(ggfortify)

autoplot(aov(groesse~sorte, blume_long))
```



#### ANOVA mit Gewässerdaten

Aus @logan2010 [S. 265]: Script an Tools/Packages des Moduls Research Methods angepasst

Medley and Clements (1998) investigated the impact of zinc contamination (and other heavy metals) on the diversity of diatom species in the USA Rocky Mountains (from Box 8.1 of Quinn and Keough (2002)). The diversity of diatoms (number of species) and degree of zinc contamination (categorized as either of high, medium, low or natural background level) were recorded from between four and six sampling stations within each of six streams known to be polluted. These data were used to test the null hypothesis that there were no differences the diversity of diatoms between different zinc levels ($$H_0 = \mu_M = \mu_L = \mu_V = \mu; \alpha_i = 0$$)

The linear effects model would be:

$$\gamma_{ij} = \mu + \alpha_i + \epsilon_{ij}$$
$$\text{diatom species diversity} = \text{overal mean} + \text{effect of zinc level} + \text{error}$$

##### Step 1

Import the Medley and Clements (1998) data set [medley.csv](13_Statistik1/data/medley.csv)
```{r}

# impact of zinc contamination (and other heavy metals) on the diversity of diatom species in the USA Rocky Mountains

medley <- read_delim("13_Statistik1/data/medley.csv", ",")

```


##### Step 2

Reorganize the levels of the categorical factor into a more logical order

```{r}

# Reorganize the levels of the categorical factor into a more logical order

medley$ZINC <- factor(medley$ZINC, levels=c("BACK", "LOW", "MED","HIGH"), ordered=T)

```

##### Step 3

Assess normality/homogeneity of variance using boxplot of species diversity against zinc group

```{r}

# Assess normality/homogeneity of variance using boxplot of species diversity against zinc group

ggplot(medley, aes(ZINC, DIVERSITY)) +
  geom_boxplot()
```

##### Step 4

Assess homogeneity of variance assumption with a table and/or plot of mean vs variance
```{r}

# Assess homogeneity of variance assumption with a table and/or plot of mean vs variance

medley_sry <- medley %>%
  group_by(ZINC) %>%
  summarise(
    mean = mean(DIVERSITY),
    var = var(DIVERSITY),
    sd = sd(DIVERSITY)
  )

ggplot(medley_sry, aes(mean,var)) +
  geom_point()
```



### Libraries

```{r code=readLines('00_Admin/get_chapter_references.R'), echo=F, eval=T,purl = F, results="asis"}
```

