---
title: Introduction to raster datasets
output: distill::distill_article
categories:
- RaumAn4
draft: false
description: |
  How we can import a raster dataset using R
author:
  - name: Patrick Laube
  - name: Nils Ratnaweera
  - name: Nikolaos Bakogiannis
---

One of the important aspects of the upcoming exercise (Multi-Criteria Evaluation (MCE)) is the use and manipulation of **raster datasets**. In `R`, two are the main packages used to handle raster data: `terra` and `raster` package. The latter is still heavily used, but inevitably is going to be replaced by the first one.  
Below we will use both of them to demonstrate how we can import a raster dataset. In the link below you can download a tif file, representing the `Digital Elevation Model` *(Digitales Höhenmodell, DHM)* of Canton Schwyz in Switzerland.

```{r, results='asis', echo = FALSE}
knitr::opts_chunk$set(echo = FALSE)

check_download <- function(files, folder, source = "", url = "https://github.com/ResearchMethods-ZHAW/datasets/raw/main/"){
  full_url <- paste0(url,folder,"/",files)
  exists_bool <- !file.exists(files)
  if(sum(exists_bool)>0){
    download.file(full_url[exists_bool],files[exists_bool])
    }
  cat(paste0("- [",files,"](",full_url,") ",source),sep = "\n")
}

check_download(c("dhm25m.tif"),"rauman")
```

### Importing raster data with `raster`

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(raster)

raster_schwyz <- raster("dhm25m.tif") 
raster_schwyz # retrieving raster info
```

### Importing raster data with `terra`

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(terra)

terra_schwyz <- terra::rast("dhm25m.tif")
terra_schwyz # retrieving raster info
```


### Plotting raster data

To plot a raster dataset, we can simply use either of the following `plot()` functions:  

- `terra::plot()`
- `raster::plot()`

```{r echo=TRUE, message=FALSE, warning=FALSE}
terra::plot(terra_schwyz)

# raster::plot(raster_schwyz)
```

